---
title: "Software tutorials"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Software tutorials}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

In this tutorial, we demonstrate basic use of the `PGInference` package. 

First we load relevant packages:
```{r,message = FALSE,warning=FALSE}
require(PGInference)
require(lattice)
require(genlasso)
require(ggplot2)
require(latex2exp)
```

We first construct a signal that is piecewise constant on an eight by eight grid graph, with three true connected components of $\beta$, which take on values of $-3$, $0$, and $3$, respectively. 

```{r fig.align="center",  warning = FALSE, fig.height = 5, fig.width = 4}
lev1 <- 0 # mean for group 1
lev2 <- 3 # mean (absolute value) for group 2/3
sigma <- 1 # level of noise
nn <- 8 # grid size
Dmat <- graph2D_Dmat(nn^2) # generate D matrix for the 2D fused lasso
# Create the underlying signal
A <- matrix(lev1, ncol=nn, nrow = nn)
A[1:round(nn/3),1:round(nn/3)] <- 1*lev2
A[(nn-2):(nn),(nn-2):(nn)] <- -1*lev2

A_true_value_plot <- data.frame(expand.grid(c(1:8),c(1:8)), z = c(A))
ggplot(A_true_value_plot, aes(x=Var1, y=Var2, fill = z)) +
  geom_tile() +
  xlab("") + ylab("") +
  ggtitle("True signal")+
  scale_fill_distiller(palette = "PiYG")+
  theme_bw()+
  theme(plot.title=element_text(size=15,hjust = 0.5),
        axis.text.y=element_text(size=15),
        axis.text.x=element_text(size=15),
        legend.position="bottom",
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        axis.title=element_text(size=15),
        strip.text.y = element_text(size=15,hjust=0,vjust = 1,angle=180,face="bold"))+
  scale_x_discrete(expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0)) +
  annotate("text",x = 2 ,y = 2, label = TeX("$C_1$"),parse=TRUE,size=6)+
  annotate("text",x = 7 ,y = 7, label = TeX("$C_3$"),parse=TRUE,size=6)+
  annotate("text",x = 4.5 ,y = 4.5, label = TeX("$C_2$"),parse=TRUE,size=6)+
  guides(fill = guide_colourbar(barwidth = 10, barheight = 1))
```

We generated data on this grid according to the model
$Y_i = \beta_i + \epsilon_i, \epsilon_i \overset{i.i.d.}{\sim} N(0, \sigma^2), \; i = 1,\ldots, 64$ and display the observation below.
```{r fig.align="center",  fig.height = 5, fig.width = 4, warning = FALSE}
set.seed(2005)
A.noisy <- A + rnorm(nn^2,mean=0,sd=sigma)
y0 <- do.call(c,lapply(1:nrow(A.noisy),function(irow)A.noisy[irow,]))
A_noisy_value_plot <- data.frame(expand.grid(c(1:8),c(1:8)), z = c(A.noisy))
ggplot(A_noisy_value_plot, aes(Var1, Var2, fill = z)) +
  geom_tile() +
  xlab("") + ylab("") +
  ggtitle("Noisy realization")+
  scale_fill_distiller(palette = "PiYG")+
  theme_bw()+
  theme(plot.title=element_text(size=15,hjust = 0.5),
        axis.text.y=element_text(size=15),
        axis.text.x=element_text(size=15),
        legend.position="bottom",
        legend.title = element_blank(),
        # axis.ticks.y=element_blank(),
        legend.text = element_text(size=15),
        axis.title=element_text(size=15),
        strip.text.y = element_text(size=15,hjust=0,vjust = 1,angle=180,face="bold"))+
  scale_x_discrete(expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0))+
  guides(fill = guide_colourbar(barwidth = 10, barheight = 1))
```

### The graph fused lasso estimate
In the code below, we use the `fusedlasso` function in the `genlasso` package to explore the estimate signals using the graph fused lasso. We display the estimate $\hat{\beta}$ after 13 steps in the figure below. Note that on this particular dataset, $K=13$ corresponds to $\lambda\in(4.73,3.42)$.
```{r fig.align="center",  fig.height = 5, fig.width = 4, warning = FALSE}
K_init = 13
complete_sol <- fusedlasso(y=y0,D=Dmat,maxsteps=50)
cat("lambda range", complete_sol$lambda[K_init-1],complete_sol$lambda[K_init],"\n")
beta_hat <- complete_sol$beta[,13]
A_estimate_value_plot <- data.frame(expand.grid(c(1:8),c(1:8)), z = c(beta_hat))
ggplot(A_estimate_value_plot, aes(Var1, Var2, fill = z)) +
  geom_tile() +
  xlab("") + ylab("") +
  ggtitle("Estimated signal with K=13")+
  scale_fill_distiller(palette = "PiYG")+
  theme_bw()+
  theme(plot.title=element_text(size=15,hjust = 0.5),
        axis.text.y=element_text(size=15),
        axis.text.x=element_text(size=15),
        legend.position="bottom",
        legend.title = element_blank(),
        legend.text = element_text(size=15),
        axis.title=element_text(size=15),
        strip.text.y = element_text(size=15,hjust=0,vjust = 1,angle=180,face="bold"))+
  scale_x_discrete(expand = c(0,0)) +
  scale_y_discrete(expand = c(0,0))+
    annotate("text",x = 2 ,y = 2, label = TeX("$\\hat{C}_1$"),parse=TRUE,size=6)+
  annotate("text",x = 7 ,y = 7, label = TeX("$\\hat{C}_3$"),parse=TRUE,size=6)+
  annotate("text",x = 4.5 ,y = 4.5, label = TeX("$\\hat{C}_2$"),parse=TRUE,size=6)+
  guides(fill = guide_colourbar(barwidth = 10, barheight = 1))
```


### Powerful graph fused lasso inference 
In this section, we demonstrate how to use our software to obtain $p$-values and confidence intervals for the difference in means of connected components 1 and 3.

Recall that for a given $\hat\tau_j$, our proposed $p$-value for testing the null hypothesis $H_0: \nu^\top \beta = 0$ can be expressed as 
$\mathbb{P}\left(|\phi| \geq |\nu^\top y| \;\middle|\; \phi \in \mathcal{S}_{C_1,C_2},$
where $\mathcal{S}_{C_1,C_2}$ is the set of values $\phi$ such that performing the graph fused lasso on the perturbed dataset $y'(\phi)$ leads to a set of connected components that include $\hat{C}_1$ and $\hat{C}_2$ (see details in the [technical details section](https://yiqunchen.github.io/PGInference/articles/technical_details.html)). 

The code below demonstrates how to use the function `fusedlasso_inf`, which estimates the connected components with $K$-step dual-path algorithm *and* performs inference on the specified two connected components based on the input data. After performing inference with $K=13$ in this example, we use the `summary` method to get a summary (in the form of a data frame) of the results.

```{r}
result_demo <- fusedlasso_inf(y=y0, D=Dmat, c1=1, c2=3, method="K", sigma=sigma, K=K_init,
compute_ci = TRUE) 
summary(result_demo)
```

We see that because $p_{Hyun}$ conditions on too much information, its power is extremely low, so that the null hypothesis cannot be rejected. By contrast, $p_{C_1,C_2}$ can easily this null hypothesis.


### Miscellaneous functions
In addition to the aforementioned key functions, the package also includes the following helper functions: `construct_v` and `estimate_spike_by_firing_rate`.

#### construct_v
The function `construct_v` generates the contrast vector $\nu$ for the null hypothesis $H_0:\nu^\top c = 0$. We will demonstrate below that $\nu^\top c$ is indeed the change in calcium in the neighborhood of a given spike $\hat\tau_j$. We consider the case where there is no spike, i.e., the calcium is decaying exponentially, and we would expect $\nu^\top c=0$. In the example below, we set $c_1 = 3$, $c_t = 0.98 c_{t-1}, \; t=2,\ldots, 50$, $\hat\tau_j = 20$, and $h=5$.

```{r}
gam <- 0.98
T_length <- 50
calcium <- rep(3 * gam ^(0:(T_length - 1)))
thj <- 20
h <- 5
```
We see that $\nu^\top c$ indeed evaluates to 0, up to machine precision.

#### estimate_spike_by_firing_rate

We have taken $\lambda$ to be known in this tutorial. Choosing an appropriate $\lambda$ value is critical to the quality of estimated spikes. The function  `estimate_spike_by_firing_rate` allows us to specify a target firing rate, and then to find the value of $\lambda$ that yields a number of estimated spikes that is close to that target firing rate.

In the example below, we simulate from an AR-1 model with firing rate 0.05, i.e., we expect to have 5 spikes per 100 timesteps on average.
```{r}
gam <- 0.98
T_length <- 1000
firing_rate <- 0.05
target_rate <- 0.05
sigma <- 0.1

```

